Index measures for oak decline severity using phenotypic descriptors

### 2020-03-04 manuscript/manuscript.Rmd
added reference to the need for tree size standardisation in intro. Spell check

The visual, manual assignment of decline status and severity in the field is variable and error prone due to the many factors contributing to oak decline.
Doing this accurately requires much expertise, experience, and training [@ferretti1997].

@@

## Phenotypic descriptor preparation {#method-decriptor-preparation}

R version `r str_c(version$major,version$minor,sep ='.')` was used for all parsing, preparation and subsequent analysis of phenotypic data [@R2019].
All raw data, data preparation routines and analysis code used to generate this manuscript are available at https://github.com/jasenfinch/Index_measures_for_oak_decline_severity_using_phenotypic_descriptors.

@@

To account for age, size and site specific tree difference, an adjustment of site means to the overall mean was applied to each of the size descriptors (DBH, lower crown height, timber height, total height, crown radius). 
An example of this adjustment is shown for DBH in Supplementary Figure 1 and the adjustment factors for each of the descriptors at each site is shown in Supplementary Table 2.  
All composite descriptors utilising these adjusted descriptors were recalculated after adjustment.

@@

## Key descriptors contributing to the decline indexes {#results-descriptor-contirbutions}

@@

The complex descriptor interactions within the index models ($M_{PDI}$, $M_{DAI}$) were investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as model response surface plots, to allow interpretation of descriptor interactions between the indexes. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for the key phenotypic descriptors described in Section \@ref(results-descriptor-contirbutions) that contribute highly to the PDI and DAI.

@@

Scaling this technique to enable computation of decline indexes for any tree sampled in the UK would require index models trained using a substantially larger data set that incorporates the full descriptive diversity of oak health in the population as whole.
Scaling to this level would also require an appropriate method of tree size correction that could be applied on an individual tree basis.
A potential solution to this would be to correct trees based on estimated age from DBH as proposed by @white1998.
These indexes could have great utility in long term forest health monitoring by providing consistent scores which can be compared over time to inform both forestry management and research [@denman2010].
For instance, the increase of the PDI values of trees within a woodland, beyond what is considered the healthy region of the PDI scale, could indicate that mangement intervention is required.

@@

```{r check,comment='',cache=FALSE}
manuscript_check(checks = c('word count')) # ,'spelling'
```

>>>>

The visual, manual assignment of decline status and severity in the field is variable and error prone due to the many factors contributing to oak decline.
For instance, accurate and reproducible crown condition assessments often require rigorous quality assurance steps that include training and regular calibration among surveyors [@ferretti1997; @redfern2004].
Additionally, differences in woodland age and individual tree sizes at different geographical locations can add difficulty to objectively comparing health status between sites.
These factors require standardisation to reduce the potential confounding comparisons for tree health status [@zarnoch2004].

@@

## Phenotypic descriptor preparation {#method-descriptor-preparation}

R version `r str_c(version$major,version$minor,sep ='.')` was used for all parsing, preparation and subsequent analysis of phenotypic data [@R2020].
All raw data, data preparation routines and analysis code used to generate this manuscript are available at https://github.com/jasenfinch/Index_measures_for_oak_decline_severity_using_phenotypic_descriptors.

@@

To account for age, size and site specific tree differences, the tree size descriptors (DBH, lower crown height, timber height, total height, crown radius) were standardised using an adjustment of site means to the overall mean. 
An example of this adjustment is shown for DBH in Supplementary Figure 1 and the adjustment factors for each of the descriptors at each site is shown in Supplementary Table 2.  
All composite descriptors utilising these adjusted descriptors were recalculated after adjustment.

@@

## Key descriptors contributing to the decline indexes {#results-descriptor-contributions}

@@

The complex descriptor interactions within the index models ($M_{PDI}$, $M_{DAI}$) were investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as model response surface plots, to allow interpretation of descriptor interactions between the indexes. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for the key phenotypic descriptors described in Section \@ref(results-descriptor-contributions) that contribute highly to the PDI and DAI.

@@

Scaling this technique to enable computation of decline indexes for any tree sampled in the UK would require index models trained using a substantially larger data set that incorporates the full descriptive diversity of oak health in the population as whole.
Scaling to this level would also require an appropriate method of tree size correction that could be applied on an individual tree basis.
A potential solution to this would be to correct trees based on estimated age from DBH as proposed by @white1998.
These indexes could have great utility in long term forest health monitoring by providing consistent scores which can be compared over time to inform both forestry management and research [@denman2010].
For instance, the increase of the PDI values of trees within a woodland, beyond what is considered the healthy region of the PDI scale, could indicate that management intervention is required.

@@

```{r check,comment='',cache=FALSE}
# manuscript_check(checks = c('word count')) # ,'spelling'
```
